3D Regression Heat Map Analysis of Population Study Data

### 2015-06-22 draft.tex
renamed cube figure

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/cube}
 \caption{
 (a) Overview visualization using a 2D heat map for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The descriptive metrics extracted from the regression model are mapped to color saturation (for $R^2$, saturated color shows a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heat map visualizations.
 %%
 These represent the slices in our \emph{3D Regression Heat Map}.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heat map.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{3D Regression Heat Map} analysis we are interested in the quality-of-fit of the resulting model.
%%
This allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment.
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{3D Regression Heat Map} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{3D Regression Heat Map} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same 2D heat map by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Cube}a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the heat map with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
%%
$AIC$ values are not restricted to a normalized scale.
%%
They have to be normalized in order to map it onto color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
Small $AIC$ values indicate a good model.
%%
Hence, we inverted the transfer function mapping it to color, mapping low $AIC$ features to saturated colors.
%%
}
\design{
%%
To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
%%
}

\paragraph{3D View.}
%%
Introducing $Z$ creates a 3D heat map (Fig.~\ref{fig:Cube}b).
%%
\add{The selected metric (per default set to $R^2$)} of each heat map entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
%%
\magenta{Object size is not used to encode information because it would result in a cluttered view.}
%%
\design{
%%
Epidemiologists with background on statistics argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
Therefore, we decided to map metrics of different regression methods on distinct colors (i.e., orange for linear regression and blue for logistic regression).
%%
}
%%
%The visualization of descriptive metrics derived from different 3D Regression Heat Maps (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
%Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., orange for linear regression and blue for logistic regression).
%%
Thus, the visualization can be easily extended using other regression types.
%%
For \emph{3D Regression Heat Maps} with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
\magenta{
As mentioned previously, the dimension reduction using the CFS algorithm potentially removes features, which are of interest for the domain expert.
%%
To tackle this problem, the $z$ dimension of the visualization contains all features of the data set, allowing to assess their influence on the models.
%%
The $x$ and $y$ dimensions are restricted to the features extracted from the CFS algorithm.
}
%%
\\\\
%%
Our goal is to create an overview visualization for a data set.
%%
We also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{3D Regression Heat Map}.

@@

\paragraph{Usage of a Regression Prism for Information Reduction.}
%%
Figure~\ref{fig:Cube} shows that all values are mirrored along the diagonal of the 2D heat map matrix.
%%
This is due to the symmetry of basic regression operators.
%%
$Z \sim X + Y$ produces the same result as $Z \sim Y + X$.
%%
Therefore, we can discard half of the results to reduce visual clutter and repetition, yielding a \emph{Regression Prism}.
%%
This opens up space for displaying additional information.
%%
Along the diagonal, $X$ and $Y$ assume the same feature, $Z \sim X + Y$ turns into $Z \sim X$ because the regression automatically ignores doublings.
%%
The diagonal therefore acts as reference on how strong the correlation for the given row (or column) feature is.

>>>>

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/heat_map}
 \caption{
 (a) Overview visualization using a 2D heat map of the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The descriptive metrics extracted from the regression model are mapped to color saturation (for $R^2$, saturated color shows a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ 2D heat map visualizations.
 %%
 These represent the slices in our \emph{3D Regression Heat Map}.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Heat_Map}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a 2D heat map.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{3D Regression Heat Map} analysis we are interested in the quality-of-fit of the resulting model.
%%
This allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{, adjusted $R^2$ and AIC} value are metrics allowing for this kind of assessment.
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{3D Regression Heat Map} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{3D Regression Heat Map} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same 2D heat map by translating the $R^2$ values to \emph{color saturation} (Fig.~\ref{fig:Heat_Map}a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the heat map with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
%%
$AIC$ values are not restricted to a normalized scale.
%%
They have to be normalized in order to map it onto color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
Small $AIC$ values indicate a good model.
%%
Hence, we inverted the transfer function mapping it to color, mapping low $AIC$ features to saturated colors.
%%
}
\design{
%%
To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Heat Map} is set per default to show $R^2$ values.
%%
}

\paragraph{3D View.}
%%
Introducing $Z$ creates a 3D heat map (Fig.~\ref{fig:Heat_Map}b).
%%
\add{The selected metric (per default set to $R^2$)} of each heat map entry (\emph{voxel}) are mapped to opacity to reduce the overlap.
%%
\magenta{Object size is not used to encode information because it would result in a cluttered view.}
%%
\design{
%%
Epidemiologists with background on statistics argued, that the visualization of descriptive metrics derived from different regression methods (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
Therefore, we decided to map metrics of different regression methods on distinct colors (i.e., orange for linear regression and blue for logistic regression).
%%
}
%%
%The visualization of descriptive metrics derived from different 3D Regression Heat Maps (e.g., $Z \sim X + Y$) is misleading, as they can be compared relatively, but not in precise numbers.
%%
%Therefore, the $R^2$ results of different regression methods are encoded using different colors (i.e., orange for linear regression and blue for logistic regression).
%%
Thus, the visualization can be easily extended using other regression types.
%%
For \emph{3D Regression Heat Maps} with a fixed target feature, e.g., $Cancer \sim X + Y + Z$, no such encodings are required and the $z$ dimension can be compared directly.
%%
\magenta{
As mentioned previously, the dimension reduction using the CFS algorithm potentially removes features, which are of interest for the domain expert.
%%
To tackle this problem, the $z$ dimension of the visualization contains all features of the data set, allowing to assess their influence on the models.
%%
The $x$ and $y$ dimensions are restricted to the features extracted from the CFS algorithm.
}
%%
\\\\
%%
Our goal is to create an overview visualization for a data set.
%%
We also want to incorporate expert knowledge into the visualization by adapting the underlying formulas.
%%
These two approaches do not exclude each other, they rather underline the difference in purpose of the chosen formula.
%%
The different analysis approaches require different starting points using the \emph{3D Regression Heat Map}.

@@

\paragraph{Usage of a Regression Prism for Information Reduction.}
%%
Figure~\ref{fig:Heat_Map} shows that all values are mirrored along the diagonal of the 2D heat map matrix.
%%
This is due to the symmetry of basic regression operators.
%%
$Z \sim X + Y$ produces the same result as $Z \sim Y + X$.
%%
Therefore, we can discard half of the results to reduce visual clutter and repetition, yielding a \emph{Regression Prism}.
%%
This opens up space for displaying additional information.
%%
Along the diagonal, $X$ and $Y$ assume the same feature, $Z \sim X + Y$ turns into $Z \sim X$ because the regression automatically ignores doublings.
%%
The diagonal therefore acts as reference on how strong the correlation for the given row (or column) feature is.
